Exploitation Techniques and Mitigations

### 2016-01-17 docs/paper.tex
Paper: Draft shell code exploits

>>>>

While this is neat and can certainly be useful to an adversary, stack smashing
also enables us to inject arbitrary code into a program. Contrary to the
previous section the target machine will execute code provided by the attacker.
This can be achieved by bending the return address into the buffer used for the
exploit. Provided instructions will be executed upon return. Shell code is a
piece of (binary) code which opens up a shell that reads and executes commands
from an attacker. Lets start this section by crafting some shell code.

\begin{nasmcode*}{autogobble,xleftmargin=1em}
    xor     eax, eax    ;Clearing eax register
    push    eax         ;Pushing NULL bytes
    push    0x68732f2f  ;Pushing //sh
    push    0x6e69622f  ;Pushing /bin
    mov     ebx, esp    ;ebx now has address of /bin//sh
    push    eax         ;Pushing NULL byte
    mov     edx, esp    ;edx now has address of NULL byte
    push    ebx         ;Pushing address of /bin//sh
    mov     ecx, esp    ;ecx now has address of address
                        ;of /bin//sh byte
    mov     al, 11      ;syscall number of execve is 11
    int     0x80        ;Make the system call
\end{nasmcode*}

This piece of assembly sets up the parameters for the \texttt{execve} system
call and than invokes to replace the currently running process with a shell.
\texttt{execve} takes three arugments, a string of the program to execute (here
\texttt{"/bin//sh"} + terminator), a list of arguments for that program and a
list of environment variables. Its system call number is 11 and it will accept
\texttt{NULL} for both lists. The double slash in the first argument is used to
prevent null bytes inside the shell code. The function which reads the shell
code may truncate it upon reading a null byte, therefore we have to work around
this without chaning the underlying semantics.

Running this code through an assembler yields binary code which can be placed
in the buffer. Finding the starting location of our buffer will be a little bit
more complicated, we cannot read it directly from the binary of the target
program so we'll examine it in a debugger.

Now we know that the buffer will be located at XXXXXXXX at runtime, but since
we got this address while running the program in a debugger it may be offset a
few bytes when run without debugger. This happens because environment variables
and meta information, like the program name, determin the stack starting
position (they are placed right after the stack). Hence we may not directly hit
the first instruction of our shell code right away, but since the buffer is
bigger than the actual payload we can improve our odds by prefixing the shell
code with \texttt{NOP} instructions. As long as the return address points
somewhere into this sequence of \texttt{NOP}s the CPU will \emph{slide} to the
next instruction. Therefore this is known as a \textit{\texttt{NOP} Sled}. We
append some arbitrary data to the shell code as offset to overwrite the return
address.

\begin{figure}[H]
    \centering
    \input{gfx/shellcode.tikz}
    \caption{Putting the payload together}
    \label{fig:shell_code}
\end{figure}

\newpage
